\thetitle

### 2021-02-09 writing/reviews/pldi-rebuttal-notes.md

Response notes format: indented after ⟶ or TODO.
⟶ is a drafted response, TODO records notes about what to respond to or how.

PLDI 2021 Paper #66 Reviews and Comments
===========================================================================
Paper #66 How to Evaluate Blame for Gradual Types

Review #66A
===========================================================================

Overall merit
-------------
A. I will champion accepting this paper.

Reviewer expertise
------------------
Z. **Some familiarity** = "I have a passing knowledge of the topic(s) but
   do not follow the relevant literature"

Paper summary
-------------
This paper presents a methodology for evaluating blame assignment strategies
based on the notion of "the rational programmer". The paper develops the
methodology in detail and then then uses it to evaluate three different blame
strategies: "Transient", "Natural" and "Erasure".

In gradual typing, a blame assignment is used to show where the program went
wrong. Importantly, at least on the theory side, a well-typed component cannot
be blamed. The authors argue that such blame assignment has received a lot of
academic attention, yet industrial implementations of programming languages with
gradual typing completely forego blame. Why is this? Is it because blame is not
useful in practice? The paper aims to answer this question by applying the
developed methodology to three strategies: The "Natural" blame strategy relies
on proxies to perform additional type checks at run-time. The "Transient" blame
strategy inlines these checks and does not perform "deep" type checks. The
"Erasure" blame "strategy" performs no checks: it only fails if execution of the
program fails.

The contribution of the paper is the methodology for evaluating blame (and
secondary its application to the three strategies.) The idea is to design an
algorithm that behaves as a programmer would debug a gradually typed program
that crashes at run-time with a type error: by incrementally adding types to the
untyped parts of the program until the source of the type error is revealed. In
this way, the question is: Does the blame assignment (or lack thereof in the
case of "Erasure") help guide the programmer? How many "steps" must he take to
find the defect and can he find it at all?

Trying to summarize the methodology:

- Take a collection of fully typed programs.
- Consider mix-typed versions of those programs.
- Develop mutators (like in mutation testing) that inject type errors.
- Emulate the rational programmer who follows the "trail" in the spirit of
  Figure 5.
  - At each step the programmer decides on an untyped component and then types
    it, reruns the program, and either sees a type error or decides to type
    another untyped component.
- Take into account how the "next untyped component" is selected based on the
  specific system at hand. 
- Experimentally evaluate whether the above method is able to find the fault and
  through how many steps.

Comments for author
-------------------
This paper is exceptionally well-written and was a joy to read. Every time I had
a concern the paper seemed to anticipate it and provide useful commentary.

The paper does a very good job at describing the challenges of how to evaluate
blame assignment strategies and then sets out to develop an automated evaluation
methodology based on the notion of "the rational programmer". The paper
carefully lays out the foundation for the generation of buggy mix-type programs,
the actions and variants ("modes") of the rational programmer which can depend
on the specific blame system, and finally how to sample the "experimental
space". The paper also includes a clear section on threats to validity.

I think it is important to have a methodology to evaluate the usefulness of
blame, and in lieu of large-scale and costly user studies, I think this paper
provides a very well thought out automatic approach (with the natural
limitations that such an approach must inevitably have.)

Points for:

- Addresses a problem that has not been addressed before.
- Clear, concise, and meaningful methodology.
- Clear description of the threats to validity.
- Very well-written paper.

Points against:
- none

# Questions

- Could you comment a bit more on what makes a mutator "good"? (other than
  experimentally seeing that it is good.) Perhaps, what would be a really bad
  mutator for evaluating blame assignment?

  ⟶ Abstractly, a good mutator reliably generates bugs that are detectable by the type system and by all three semantics' runtime type checks; ideally the bugs also manifest during the interaction of multiple components.
    We rely on the experimental check of our mutators since this criteria is difficult or impossible to decide otherwise.
    That said, we can reason that some mutators are clearly unlikely or impossible to be good.
	...
	We will extend the discussion at the start of section 4 to clarify this point.

  TODO: Rewrite response along these lines...
    A common mutator in the literature is swapping constants of the same type, but this is not a good mutator for us because type system can't catch it.
	Example of an obviously bad mutator: replacing any string with some pre-defined non-empty string (e.g. "foobar") would be a bad mutator.
	Examples of mutators we cut: `wrap-cond` (tricks occurrence typing by hiding the result of conditionals to produce a type error, but never a dynamic error), `add-extra-method`, swapping `car` and `cdr`, `set!`ing an ID to itself (only causes a type error), swapping mutable and immutable hashes+vectors.
	# This answer should serve to justify our mutators and the challenge of making good mutators more broadly, to answer reviewer D's concerns about the delta of making mutators.

- I don't understand where the 16,800 number comes from. Is it an exhaustive
  enumeration of all possible mutations? (Or the number of mutants constructed
  in X amount of time?)
  ⟶ The former, we will clarify in the text.

- How large are the programs in the benchmark suite?
  ⟶ The program sizes without type annotations, as reported by Racket's `syntax-sloc` tool, are below. We will add this information to the table in Figure 2.
    benchmark | LOC | components
	----------|-----|-----------
	acquire | 1941 | 9
	gregor | 2336 | 13
	kcfa | 328 | 7
	quadT | 7396 | 14
	quadU | 7282 | 14
	snake | 182 | 8
	suffixtree | 1500 | 6
	synth | 871 | 10
	take5 | 465 | 8
	tetris | 280 | 9

# Minor Comments

- L38: Should there have been a colon? Is the text "This contrast... " and
  challenge quote from [3]? Who is saying what here...
  ⟶ Thanks, we will reword the text from "This contrast" to clarify that this our research question, it is not coming from [3].

- L104: I was confused by the word "Next". Maybe instead: We now ..."
  ⟶ Thanks, we will change the wording.

- L268: Grammar issue.
  ⟶ Thanks, we will fix this.

- L314: Floating period.
  ⟶ Thanks, we will fix this.

- L441: Add line numbers (or other metric of size).
  ⟶ See above, we will add this information.

- L956: 30,000 hours of compute time is 1,250 days. I assume that the CPU has
  many cores. Could you put that number?
  ⟶ Each CPU has 28 cores and 56 threads; when we last ran the experiment, it took about two weeks of real time from start to end. 
  We will add this information at the start of section 7.

- L980: Figure 5.4??
  ⟶ Thanks, this is a bad reference; it should be figure 6. We will fix it.

- L991: Figure 6: It took me some time to understand the figure. I wonder if
  there is a better way to represent the same data?
  TODO: come up with some thoughts for this figure
  ⟶ Thanks, we can see how it's difficult to read. We will consider some different presentations, such as...
    - a bunch of two-sided bar charts? Where there's one bar pair per pair of modes. Above the x-axis is % A > B, and below the axis is % B > A?
	- a bunch of stacked histograms? Where there's one stack per pair of modes. A bar has three parts: A > B, A = B, and B > A.

Review #66B
===========================================================================

Overall merit
-------------
B. I support accepting this paper but will not champion it.

Reviewer expertise
------------------
Y. **Knowledgeable** = "I follow the literature on this paper's topic(s)
   but may have missed some relevant developments"

Paper summary
-------------
The paper proposes an evaluation method for blame assignment
strategies and evaluates three different strategies with the
method. While many theoretical gradual typing systems have adopted
blame, blame has been neither shown to be practically useful nor
applied to real-world languages because the systematic evaluation of blame
is difficult. The authors introduce a notion of rational programmers
and simulate their behaviors to evaluate blame. A rational programmer
is assumed to iteratively add type annotations to an untyped portion
of his or her program according to the feedback given by a run-time
system. This behavior can be simulated even without real human
programmers. If a run-time system leads a programmer to the correct
locations of a bug, its blame strategy can be considered more useful
than a strategy that cannot do so. The paper compares the Natural,
Transient, and Erasure strategies. The result shows that both Natural
and Transient outperform Erasure, which does not assign any blame,
and, therefore, implies usefulness of blame assignments.

  TODO: This seems to imply that we are proposing or assuming the way that real "rational programmers" would behave. Should we try to answer or clarify about that?

Comments for author
-------------------
This paper tackles a long-standing difficult problem, which is the
problem of evaluating blame. Many existing gradual typing papers
provide blame systems, but because the papers do not evaluate blame
systems, it is not obvious whether blame itself is indeed helpful for
debugging. Among many theoretical gradual type systems adopting blame
assignments, no paper discussed the evaluation of blame systems. It is an
important problem because, without any evaluation, theoretical aspects
like gradual guarantee and the blame theorem may not be applicable to
practical systems combining static and dynamic typing.

This paper shows that blame assignments do help programmers, and it
may serve as a first step to close the gap between the theory and
the practice of gradual typing.

The experiment results show that there are debugging scenarios where
Transient is more useful than Natural. Does it happen simply because
of luck? Or, does it imply that Transient is more suitable than
Natural for a particular kind of bugs for some reason? Also, can there
be a new blame strategy that outperforms both Transient and Natural
all the time?

  ⟶ 
	Response idea: This is an example where the experiment seems to contradict the theory. 
	The models don't include all of the features in the benchmarks. <enumerate them here> It's possible that the Natural, Transient theories don't scale as expected to rich PL features.
	Make analogy to physics: theory makes predictions, we have found failures of the predictions. We have two courses of action: report the inconsistency to re-eval the theory, and do more experiments.
	It's also possible we have a bug in TR, or in Transient implementation.

    Later: TODO: Scenarios where Transient is more useful than Natural imply that Transient locates the bug while Natural gets stuck.
	      check what the reasons for those scenarios are;
	      - the % is too high to all be because of the classes bug,
		  - how else could Natural get stuck if Transient isn't? It can't be the bad exn cases, because then Transient should be stuck as well because Transient shouldn't be able to get blame in those cases.

Theoretical gradual type systems usually allow programmers to control
the precision of type annotations in a fine-grained way with the Dyn
type rather than distinguishing typed and untyped modules. Would the
same evaluation method be able to directly appliable to blame in such systems?

  TODO: The answer is yes, but this needs more thought to frame the answer the right way.
  ⟶ Our evaluation framework should be directly applicable to blame in gradual typing systems using Dyn.
    The details of how it applies depends on the granularity of the blame information offered by the system -- in other words, what the system considers components.
	If components are modules, as in Typed Racket, then nothing changes.
	Systems with Dyn more likely treat top-level definitions as components, in which case definitions are individually typed (the whole definition is annotated) rather than modules.

Please explain what impedance mismatches are in the paper. Is it
simply type mismatches?

  ⟶ Impedance mismatches are mismatches between a piece of code's true type and its annotation.
    For instance, the code `(λ (x) (+ x 42))` has type `Number -> Number` but might be annotated as `Number -> String`, causing an impedance mismatch.
    We will adjust the text to clarify or remove this terminology.

# Strengths

* The paper is well-organized.
* It proposes a novel solution to a long-standing difficult problem.
* Threats to validity are discussed in detail.

# Minor comments

The paper states that the effort distribution of the random mode
follows a normal distribution, as expected. Please explain why a
normal distribution is a reasonable expectation. When there are $n$
untyped components, the random mode randomly selects one of them, and
only one of them is a buggy component. Therefore, in the first
iteration, the random mode reaches the end of the trail with the
probability of $1/n$. In the second iteration, there are $(n-1)$
untyped components and the probability of a failure at the first
iteration is $(n-1)/n$, so the probability of effort $2$ is $1/n$
again. In this way, it seems to be that the probability of effort $i$
is $1/n$ where $1<=i<=n$ and $n$ is the number of initially untyped
components. Each trail has a different number of untyped components,
so the resulting distribution would be the average of such
distributions. Based on this reasoning, it is unclear why the
distribution is a normal distribution. This point does not affect the
contribution of the paper, but it is not clear from the paper.

  TODO: response sketch: you're right, we misspoke, this is not the normal distribution; our reasoning for why the normal-ish shape of distribution we see makes sense is that given a fixed trail the probability at each step of the trail ending is 1/n, but we randomly select trails so we're multiplying the random distribution of trail lengths by that 1/n one? Not clear on the details of this reasoning.
  ⟶

* page 9 line 980: Figure 5.4 -> Figure 6
  ⟶ Thanks, we will fix this.
* page 10 Figure 7: Null -> Random
  ⟶ Thanks, we will fix this.
* page 11 line 1103: ( 770 -> (770
  ⟶ Thanks, we will fix this.

Review #66C
===========================================================================

Overall merit
-------------
C. I would not accept this paper but will not argue strongly against
   accepting it.

Reviewer expertise
------------------
Z. **Some familiarity** = "I have a passing knowledge of the topic(s) but
   do not follow the relevant literature"

Paper summary
-------------
This paper develops an approach for empirically evaluating the effectiveness of blame assignment in gradually typed languages. Roughly speaking, the idea is to model the steps taken by a "rational programmer" who systematically debugs errors in a gradually typed programming. The question is whether different blame assignment strategies are able to help this rational programmer find the "true" source of errors faster, requiring fewer steps.

To evaluate this approach, the paper also presents empirical results. Starting with a small corpus of programs, the authors use mutations to systematically (but synthetically) introduce type errors -- e.g., one mutation replaces a constants with another of a different type, and another mutation swaps method identifiers. They then conduct experiments to quantify the number of steps required by a rational programmer for several blame assignment schemes: Natural (used in Typed Racket, which tracks at most one location for blame), Transient (used in Reticulated Python, which tracks a set of locations), and Erasure (which does not actually assign blame, and is widely used in industry).

  TODO: do we need to clarify that effort is a secondary aspect of the evaluation, and more important is whether blame leads to the bug at all?

Unsurprisingly, the evaluation shows that blame is effective, when used by a rational programmer. The results are inconclusive, but some trends are clear. Most significant, all of the schemes that assign blame are more useful than Erasure. And Natural blame more often more useful than Transient blame.

Comments for author
-------------------
This is a neat paper that develops an approach for evaluating gradual typing and blame assignment. Thanks for submitting it to PLDI '21! The results are presented in an extremely clear way that draws out the key insights without compromising on technical detail. The notion of a rational programmer -- loosely modeled on rational agents from economic theory -- is a neat conceptual device. And the empirical approach does seem like the right way to address the question of the usefulness of blame.

But while there is a lot to like about the paper, it also has some shortcomings. The suite of programs used for the evaluation is fairly small (c.f., the SIGPLAN checklist).
  TODO: respond to this ^, saying that we use the standard suite of benchmarks for evaluating GT -- point to Ben's POPL paper and others, and we use all of the largest benchmarks in the suite
The faults are introduced using synthetic mutations that may or may not correspond to the kinds of errors that arise in practice. This can be seen from the results in Figure 7 -- the lengths of the "trails" is quite small.
  TODO: unclear that syntheticness relates to trail length at all, and synthetic bugs are the only option here -- there is no corpus of documented GT bugs in the wild anywhere. Indeed, most type errors never even make it into code repositories. Note that we are the first to evaluate blame in this setting.
While the study is mostly well done, it only considers two blame assignment strategies. For example, it would be interesting to consider the "omniscient programmer" to get a bound on the value of blame. But this is not done.
  TODO: We don't understand what is meant by "omniscient programmer", but our best understanding is the ideal oracle strategy that always picks the right module right away and thus always finds the bug in one step. We didn't show this because it's trivial. We do have other baselines to understand blame's performance: the random mode, and the exceptions modes for each of the semantics.
And unfortunately the results are inconclusive -- Section 8 says "our results call for a deeper understanding of the two models for blame".
  TODO: This is what makes our results surprising: they challenge the theoretical conclusions that Natural is superior to Transient, suggesting that the theory is not good enough to understand the pragmatics of blame in reality.
Also, the published Transient scheme seems to have a minor bug, which might affect the outcomes when fixed (Section 8.2).
  TODO: 8.2 does not indicate a bug in Transient, but rather a defect in the design of the system. ... something about a gap between Reticulated Python and 
Most if not all of these shortcomings are discussed in the paper, which is much appreciated. But they do risk undermine the value of the contributions to some degree.

To explain one comment above: I was surprised not to see a comparison against an "omniscient programmer" to establish a baseline (c.f., the SIGPLAN checklist). That is, model a blame assignment scheme in which blame could be placed on any component. And always pick the component that minimizes the length of the trail. Of course, this would be infeasible to implement in practice, but it would place an upper bound on the utility of blame.

Similarly, I wondered why the modes of the transient programmer only consider picking the first/last blame, and not other choices. Section 5.2 just says "our answer" is that there are at least "two reasonable options." Why are other choices not reasonable?
  TODO: Answer suggested by Ben: it's unspecified how one should interpret Transient blame, and there isn't time to evaluate all possible choices, so we pick the most reasonable interpretations.

Review #66D
===========================================================================

Overall merit
-------------
B. I support accepting this paper but will not champion it.

Reviewer expertise
------------------
Y. **Knowledgeable** = "I follow the literature on this paper's topic(s)
   but may have missed some relevant developments"

Paper summary
-------------
Several contract and gradual type systems provide mechanisms for
assigning blame: When encountering a run-time error signifying the
violation of a contract or type signature, which pieces of code should
be scrutinized by the programmer in seeking a fix?

In a recent POPL 2020 paper, Lazarek et al. provide the first attempt
to evaluate if and how blame assignment may help in actual debugging.
In that paper, bugs are systematically injected (via syntactic
mutation) into a suite of Racket programs, and the corresponding blame
trails are analyzed to see whether they, by writing more precise
contracts to "shift" blame from the assigned component, eventually
lead to the actual bug.

The current paper extends the POPL 2020 methodology from contracts in
Racket to gradual (or migratory) typing in Typed Racket, in several
steps:

First, to complement the "Natural" higher-order contract system of
(Typed) Racket, this work implements a "Transient" blame assignment
strategy, which unlike Natural inserts only shallow run-time checks.
(To date, Transient has appeared only in the separate design and
implementation of Reticulated Python, which prevents a head-to-head
comparison as desired in this work.)

Second, compared to POPL 2020, new mutators are defined to inject the
kinds of type-mismatch bugs that may be caught by the gradual type
system and corresponding run-time checks and blame assignment.

Third, the paper runs a large experiment on 10 benchmark programs,
inserting mutations and analyzing whether the blame trails produced by
each debugging mode --- following (a) the component blamed by Natural,
(b/c) the first or last component, respectively, blamed by Transient,
or (d) an exception raised by the underyling language --- conclude
with the actual bug. Results show that Natural, Transient-First, and
Transient-Last all produce short successful blame trails (typically of
length one or two) on most debugging tasks, and that they are all
similarly effective with no clear winner for all tasks.

### 2021-02-12 writing/challenges.tex
Minor rewordings in intro and challenges
touching up the prose

The most serious challenge for evaluating blame concerns 
the nature of this research. Unlike ordinary
work on programming languages, the question seems to call for empirical research
similar to the one in the human-computer interaction area. At the same time, a
significant evaluation result demands a large number of cases yielding a serious
amount of data. But, as~\citet{lksfd-popl-2020} already demonstrated, the way
around this dilemma is to postulate the equivalent of a {\it homo economicus\/}
for programming languages---the {\em rational programmer\/}---and to implement
this rational programmer as an algorithm.  As indicated in the introduction, in
the context of gradual typing, the rational programmer translates the
Wadler-Findler slogan into a debugging method, searching for the source of the
impedance mismatch in an incremental fashion. Measuring this simulated
programmer's behavior on a large number of inputs yields relevant 
data, similar to collecting empirical data in a human-focused discipline.

A derivative challenge is the problem of simulating the programmer's
mistakes or, more precisely, the origin of impedance mismatches. Before a
programmer can try to find mistakes, mistakes have to be made.
Specifically, a necessary step to test the rational programmer is to find
a collection of  type mistakes in  mix-typed programs that is highly
representantive of both mistakes and programs ``in the wild.''
Unfortunately no such collection exists.  There is a good reason for that;
the kind of mistakes the rational programmer targets are easily detected
by unit tests and, even if their sources are hard to find, these mistakes do not make it
into code repositories with appropriate commit messages.  In other words, the problem is how to generate a
sufficiently large and representantive corpus of type mistakes for a meaningful
experiment.  Following~\citet{lksfd-popl-2020} again, the solution is to
use mutation analysis~\cite{lipton1971fault, demillo1978hints,
jia2011analysis} to create mistakes in a synthetic manner. This process
yields a large number of buggy programs whose mixed-typed versions are
suitable subjects for our experiment.  The mutation analysis of Lazarek et
al, is, however, useless for creating impedance mismatches.

While mutation analysis traditionally aims to inject bugs that challenge
extensive test suites, the ones yielding ill-typed mutants are usually deemed
\emph{incompetent}. As a matter of fact, mutation analysis frameworks are
fine-tuned to avoid them---yet it is precisely those that are needed for
evaluating blame assignment strategies. The mutators used here are newly designed,
inspired by the authors extensive experience with programming in Typed Racket.
They cause type errors that cover a wide range of Typed Racket's type system and moreover
reliably cause runtime exceptions in mixed-typed versions of mutants.

The second big challenge is the disparity of gradually
typed languages. While the approaches are well understood in theory,
they come with vastly different type systems, accommodate seemingly
incompatible language features, and apply to radically different coding
styles in the untyped world.  To eliminate the effect of these differences 
on its results, the application of an evaluation method to different
gradual typing systems must use the same test programs and the same type
system. Otherwise, results for different gradual typing systems 
do not share a common baseline and, therefore, do not lend support to any
conclusions about whether the value of blame changes from one system to
another. For instance, the case study presented here builds on
Greenman and Felleisen's Transient variant for functional Typed
Racket~\cite{gf-icfp-2018} to obtain a Transient semantics that is
comparable with Typed Racket's default Natural one. This Transient
variant accommodates varied linguistic features and library calls
in a way that is faithful to the Transient
semantics~\cite{vss-popl-2017}.

Beyond the implementation issue, the method also faces the problem that
the three variants---the Natural semantics, the Transient semantics, and
the Erasure semantics---come with three radically different blame
assignment strategies.  For example, the Natural semantics assigns blame
to one component while Transient assigns blame to a set of components. The
Erasure semantics does not blame a component per se, but it comes with an
exception location and the usual stack traces.

@@

The final big challenge is that the different gradual type
systems, the number of necessary modes for the rational programmer
and the large number of mutants lead to an explosion of scenarios the
method needs to examine. Thus a practical application of the
method has no option but to carefully sample the space of scenarios while 
making sure that it still produces reproducible conclusions.

@@

These challenges expose the subtlety of evaluating blame strategies but
also suggest a recipe:

@@

\item Implement all three gradual type systems on a common
  basis (section~\ref{sec:landscape});

\item Generate a large set of programs with interesting
  type-level errors (section~\ref{sec:mutate});

\item Simulate the rational programmer on these programs with sufficient 
  modes to make meaningful comparisons between modes of the same or
    different systems
    (section~\ref{sec:rational});

\item Sample the experimental space  to obtain a computationally feasible
  but statistically sound experiment (section~\ref{sec:sample}).

>>>>

%% -----------------------------------------------------------------------------

The challenge of evaluating blame is a methodological one. Unlike most research
on programming languages, the question seems to call for empirical research
similar to that of human-computer interaction. At the same time, a significant
evaluation result demands a large number of cases yielding a serious amount of
data. As~\citet{lksfd-popl-2020} recently demonstrated, the way around this
dilemma is to postulate the equivalent of a {\it homo economicus\/} for
programming languages---the {\em rational programmer\/}---and to implement this
rational programmer as an algorithm.

In the context of gradual typing, the rational programmer translates the
Wadler-Findler slogan into a debugging method, searching for the source of the
impedance mismatch in an incremental fashion. Measuring this simulated
programmer's behavior on a large number of debugging sessions yields data that
is similar to empirical data collected in a human-focused discipline.

A derivative challenge is the problem of simulating the impedance mismatches
between typed and untyped pieces of code that programmers may accidentally create
and that the run-time checks of academic systems catch. In other words, the
experiment calls for a collection of type mistakes in mix-typed programs that is
representative of those ``in the wild.'' Unfortunately no such collection exists,
and with good reason; the kind of mistakes the rational programmer targets are
easily detected by unit tests and, even if their sources take some time to find,
these mistakes do not make it into code repositories with appropriate commit
messages.  Hence the experimenter must generate a sufficiently large and
representative corpus of type mistakes for a suitable collection of benchmark
programs.  Following~\citet{lksfd-popl-2020} again, the solution is to synthesize
mistakes using mutation analysis~\cite{lipton1971fault, demillo1978hints,
jia2011analysis}.  The mutation analysis of Lazarek et al, is, however, useless
for creating impedance mismatches.  While mutation analysis traditionally aims to
inject bugs that challenge test suites, it discards those that yield ill-typed
mutants as \emph{incompetent}.  Indeed, mutation analysis frameworks are
fine-tuned to avoid them---and yet it is precisely those that are needed for
evaluating blame assignment strategies.

Thus, an experimental analysis of blame needs a mostly new set of mutators.
Roughly speaking, the new mutators inject type errors into fully typed programs.
Applying such a mutator to any typed component produces a mutatnt component.  A
debugging scenario results from removing the types from the mutatnt. For the
design of such mutators, the authors relied on their own extensive programming
experience though not without discovering a major pitfall: some of their original
mutators systematically produced programs that immediately revealed the source of
the impedance mismatch. About two thirds of the original mutators survived this
first scrutiny testing. All of the remaining ones yield interesting debugging
sessions.

A somewhat coincidental challenge is the disparity of gradually typed
languages. While the approaches are well understood in theory, they come with
vastly different type systems, accommodate seemingly incompatible language
features, and apply to radically different coding styles in the untyped world.
To eliminate the effect of these differences on its results, the application of
an evaluation method to different gradual typing systems must use the same test
programs and the same type system. Otherwise, results for different gradual
typing systems do not share a common baseline and, therefore, do not lend support
to any conclusions about whether the value of blame changes from one system to
another. For instance, the case study presented here builds on Greenman and
Felleisen's Transient variant for functional Typed Racket~\cite{gf-icfp-2018} to
obtain a Transient semantics that is comparable with Typed Racket's default
Natural one. This Transient variant accommodates varied linguistic features and
library calls in a way that is faithful to the Transient
semantics~\cite{vss-popl-2017}.

Beyond the sheer implementation issue, the development of a method must also
address the problem that the three semantic variants---Natural, Transient, and
Erasure---come with three different blame assignment strategies.  While Natural
assigns blame to {\em one\/} component, Transient assigns blame to a sequence of
components. The Erasure semantics does not blame a component per se, but it comes
with an exception location and a stack trace.

@@

The final big challenge is the explosion of debugging scenarios that results from
the combination of mutation-based scenario generation and mode-based analysis.
Specifically, the experiment demands a large amount of computational resources
due to the multiplicative effect of mixing the different gradual type systems,
the number of debugging modes for the rational programmer, and the large number
of mutants.  A practical application of the method has no option but to sample
the space of scenarios, carefully ensuring that it still produces reproducible
conclusions.

@@

Overcoming these challenges needs a systematic approach:

@@

\item a single-language implementation of all three gradual type systems (section~\ref{sec:landscape});

\item a suitable set of mutators for the generation of a large set of debugging
  scenarios with interesting type-level errors (section~\ref{sec:mutate});

\item an algorithmic simulation of the rational programmer on these debugging
  scenarios with sufficient modes to make meaningful comparisons between modes of
  the same or different systems (section~\ref{sec:rational});

\item a careful sampling of the experimental space to obtain a computationally
  feasible and statistically sound experiment (section~\ref{sec:sample}).
